### 2015-10-04 article.tex
Structure plus first version of the abstract

\documentclass{article}

\usepackage{csquotes}

\begin{document}

\title{Functional vs. Object Oriented Programming}
\author{Lorenzo Corneo and Nikolaos-Ektoras Anestos \\
	\{corneo, anestos\}@kth.se}
\date{\today}
\maketitle

\begin{abstract}
Human resources and time consumption are very important factors in software development. Both developers and managers try to maximise their productivity, in order to increase as much as possible the profits or gain market advantage. We assume that choosing an appropriate programming paradigm is a critical factor that leads to saving of these resources.

The main object of investigation in this project are the lines of code needed to develop two different implementations of the same program and the level of reusability of existing code.

Functional programming has some fundamental benefits that increase programmer productivity over that when OOP languages. The purpose of this project is to demonstrate some of the benefits of functional programming over OOP, in order to persuade developers to switch to functional programming.
\end{abstract}

\section{Introduction}

\section{Comparison criteria}

\section{Experimentation results}

\section{Evaluation}

\section{Acknowledgement}

\section{Conclusion}

\begin{thebibliography}{9}
\bibitem{flink}
Apache Flink. 2015. Apache Flink: Scalable Batch and Stream Data Processing. [ONLINE] Available at: http://flink.apache.org. [Accessed 16 September 15].
\bibitem{rx}
ReactiveX. 2015. ReactiveX. [ONLINE] Available at: http://reactivex.io/. [Accessed 16 September 15].
\bibitem{fpoop}
R. Harrison, \enquote*{Comparing programming paradigms: an evaluation of functional and object-oriented programs}, Jul. 1996 [Online]. Available: http://ieeexplore.ieee.org/stamp/stamp.jsp?tp=\&arnumber=511273
\bibitem{githubapi}
GitHub Developer. 2015. Statistics | GitHub API. [ONLINE] Available at: https://developer.github.com/v3/repos/statistics/. [Accessed 16 September 15].
\bibitem{benkio}
https://github.com/benkio/ITC-Slides
\end{thebibliography}

\end{document}

### 2015-10-04 article.tex
Abstract and introduction

@@
-\documentclass{article}
+\documentclass{proc}
 
@@
 The purpose of this project is to demonstrate some of the benefits of functional programming over OOP, in order to persuade developers to switch to functional programming.
+
+From the evaluation of the gathered data it is possible to see that, averagely, the number of Lines Of Code in a functional programming language is smaller than in an Object Oriented Programming language.
+As a consequence, Functional Programming allows an increased productivity (fewer lines of code correspond to a time saving).
+\textbf{Then when we have data regarding the code reuse we write here.}
 \end{abstract}
+
+\smallskip
+\noindent \textbf{Keywords.} Functional Programming, Object Oriented Programming, Comparison, Productivity, Code Reuse.
 
@@
 
-\section{Comparison criteria}
+The paper presents the results of a quantitative evaluation of code implemented in both Functional Programming and Object Oriented Programming.
+The main parameters that will be analysed are the Line Of Code and the Code Reuse.
+
+The main test-bed for the productivity analysis is based on the LOC of different implementations (both FP and OOP) of very well-known algorithms.
+The language pool is composed by Erlang, Haskell, Scala, F\#, for the FP languages and by Java, C++, C\#, Python for the OOP languages.
+
+The CR is evaluated counting the difference in terms of LOC between adjacent releases of a program.
+Object of the investigation, as a case study, is the asynchronous web server Play framework, that is implemented both in Java and Scala.
+
+The results of the investigation are presented in section 3 while the evaluation of the data are discussed in section 4.
+
+The aim of the research is to show the benefits of FP over OOP with the hope to persuade software developer to adopt FP as their main programming paradigm, for a better productivity and code reuse.
 

### 2015-10-06 article.tex
First draft of literature study and bibliography added.

@@
 
-\usepackage{csquotes}
+\usepackage{natbib}
+\usepackage{url}
 
@@
 As a consequence, Functional Programming allows an increased productivity (fewer lines of code correspond to a time saving).
-\textbf{Then when we have data regarding the code reuse we write here.}
 \end{abstract}
@@
 The CR is evaluated counting the difference in terms of LOC between adjacent releases of a program.
-Object of the investigation, as a case study, is the asynchronous web server Play framework, that is implemented both in Java and Scala.
+Object of the investigation, as a case study, is the asynchronous web server Play framework\cite{play}, that is implemented both in Java and Scala.
 
-The results of the investigation are presented in section 3 while the evaluation of the data are discussed in section 4.
+In section 2, a summary of the studied literature will be presented while in section 3 the methods and the hypothesis will be discussed.
+The results of the investigation and the analysis are presented in section 5, in section 6 a discussion about the initial hypothesis and the results will be discussed.
 
@@
 
-\section{Experimentation results}
+\section{Literature study}
 
-\section{Evaluation}
+OOP is the de facto standard programming paradigm for companies and widely adopted all around the world.
+This find a confirmation in the TIOBE Index\cite{tiobe}.
+In fact, in the first positions only OOP and Web programming languages are found, while the first FP language is Scala at the 24th position with a rating of 0.781\%
 
-\section{Acknowledgement}
+Nevertheless, functional programming gurus always try to persuade imperative programmers to switch to this paradigm promising incredible advantages.
+Amongst these benefits we found the lack of assignment statements (variables never change) and side effects, which is the major source of bugs in software development\cite{fpmatter}.
+
+Furthermore, a thorough evaluation about the two programming paradigms has been performed by Harrison\cite{harrison}.
+He defined parameters to be able to measure the quality of the software, including LOC and CR.
+Then, from the domain application of the image processing, he selected a developer with similar skills both in C++ and SML (functional language of the ML family) who developed a set of 12 algorithms in both languages.
+From the analysis of the results, it turned out that there are not many differences between the two paradigms.
+
+Nonetheless, FP finds very important benefits in the application domain of the prototypes where a case study project is implemented in Haskell in 85 LOC against the 1105 of C++\cite{Haskell-vs-ada}.
+This is due to three main factors: (a) the syntax is simpler, (b) the use of higher order functions and (c) the standard list-manipulation primitives (i.e.
+map, fold, zip, ...).
+
+\section{Hypothesis}
+
+\section{Methods}
+\section{Results and Analysis}
+
+\section{Discussion}
 
@@
 
-\begin{thebibliography}{9}
-\bibitem{flink}
-Apache Flink. 2015. Apache Flink: Scalable Batch and Stream Data Processing. [ONLINE] Available at: http://flink.apache.org. [Accessed 16 September 15].
-\bibitem{rx}
-ReactiveX. 2015. ReactiveX. [ONLINE] Available at: http://reactivex.io/. [Accessed 16 September 15].
-\bibitem{fpoop}
-R.
+\bibliography{article}
+\bibliographystyle{plain}
+
+%\begin{thebibliography}{9}
+%\bibitem{tiobe}
+%http://www.tiobe.com/index.php/content/paperinfo/tpci/index.html
+%\bibitem{flink}
+%Apache Flink. 2015. Apache Flink: Scalable Batch and Stream Data Processing. [ONLINE] Available at: http://flink.apache.org. [Accessed 16 September 15].
+%\bibitem{rx}
+%ReactiveX. 2015. ReactiveX. [ONLINE] Available at: http://reactivex.io/. [Accessed 16 September 15].
+%\bibitem{fpoop}
+%R.
 Harrison, \enquote*{Comparing programming paradigms: an evaluation of functional and object-oriented programs}, Jul.
@@
 Available: http://ieeexplore.ieee.org/stamp/stamp.jsp?tp=\&arnumber=511273
-\bibitem{githubapi}
-GitHub Developer. 2015. Statistics | GitHub API. [ONLINE] Available at: https://developer.github.com/v3/repos/statistics/. [Accessed 16 September 15].
-\bibitem{benkio}
-https://github.com/benkio/ITC-Slides
-\end{thebibliography}
+%\bibitem{githubapi}
+%GitHub Developer. 2015. Statistics | GitHub API. [ONLINE] Available at: https://developer.github.com/v3/repos/statistics/. [Accessed 16 September 15].
+%\bibitem{benkio}
+%https://github.com/benkio/ITC-Slides
+%\end{thebibliography}
 

### 2015-10-06 article.tex
Hypothesis draft added.

@@
 
+Human resources and time consumption are very important factors in software development, especially for companies that want to optimise profits and gain advantages from the market.
+
+We strongly believe that, using functional programming languages, developers will be more productive and they will also produce higher quality code that will allow them to reuse much of it when releasing newer versions of their programs.
+
 \section{Methods}
+
+A quantitative method is used for this project because the authors will gather data from many sources and they will report in the paper.
+Afterward, the analysis of the gathered data will be performed so that the initial hypothesis will be either accepted or rejected.
+
+The approach used will be deductive because we started this research with our own hypothesis on the basis of existing theories, for instance, regarding the benefits of FP and the advantages that developers can gain in software production.
+
+Afterward, when the data will be acquired, the authors will perform observation and analysis of the gotten results. From that, it will be possible to accept or reject the initial hypothesis.
+
 \section{Results and Analysis}

### 2015-10-06 article.tex
Images with captions and pdf.

@@
-\documentclass{proc}
+\documentclass{article}
 
@@
 \usepackage{url}
+\usepackage{graphicx}
 
+\graphicspath{ {img/} }
 \begin{document}
@@
 
+\begin{figure}[h]
+\includegraphics[width=\textwidth]{1}
+\caption{The graph represents the distribution of both FP and OOP paradigm in different algorithms.}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=\textwidth]{2}
+\caption{The graph shows the distribution of the LOC of different languages implementation for different algorithms.}
+\end{figure}
+
+\begin{figure}[h]
+\includegraphics[width=\textwidth]{3}
+\caption{The graph shows the LOC used to solve each algorithm for the same language.}
+\end{figure}
+
 \section{Discussion}

### 2015-10-07 article.tex
First draft for result and analysis section

@@
 This find a confirmation in the TIOBE Index\cite{tiobe}.
-In fact, in the first positions only OOP and Web programming languages are found, while the first FP language is Scala at the 24th position with a rating of 0.781\%
+In fact, in the first positions only OOP and Web programming languages are found, while the first FP language is Scala at the 24th position with a very low rating of 0.781\%.
 
@@
 
-\begin{figure}[h]
-\includegraphics[width=\textwidth]{1}
-\caption{The graph represents the distribution of both FP and OOP paradigm in different algorithms.}
-\end{figure}
+The experimentation consist in static code analysis on a set of eight algorithms, namely:
+
+\begin{itemize}
+\item FASTA
+\item k-nucleotide
+\item Mandelbrot
+\item n-body
+\item pidigits
+\item regex-dna
+\item reverse-complement
+\item spectral-norm
+\end{itemize}
+
+The previous set of algorithms has been implemented in the following functional languages (both pure and not):
+
+\begin{itemize}
+\item Haskell
+\item Erlang
+\item Scala
+\item Clojure
+\item F\#
+\item Python
+\end{itemize}
+The object-oriented languages these algorithm were implemented with are:
+
+\begin{itemize}
+\item C++
+\item C\#
+\item Go
+\item Java
+\end{itemize}
+
+In the experimentation we analysed the LOC for each language per each algorithms and the results are shown in Fig.
+\ref{fig:alg}.
+In this graph it is possible to see which language is the most efficient in terms of lines of code per each single algorithm.
 
@@
 \caption{The graph shows the distribution of the LOC of different languages implementation for different algorithms.}
+\label{fig:alg}
 \end{figure}
+
+An alternative way to visualise these results is to think in terms of language and display which algorithms is the best to implement and be more productive.
+Fig \ref{fig:lang} shows this alternative representation.
 
@@
 \caption{The graph shows the LOC used to solve each algorithm for the same language.}
+\label{fig:lang}
 \end{figure}
+
+The above graphs have the disadvantages to present too much information so that the readers may have trouble understanding which paradigm is more efficient.
+For this reason we synthesised the gathered data in a graph that shows only the average LOC for each paradigm per each algorithm.
+This representation of the data is shown in Fig.
+\ref{fig:fpvsoop}.
+
+\begin{figure}[h]
+\includegraphics[width=\textwidth]{1}
+\caption{The graph represents the distribution of both FP and OOP paradigm in different algorithms.}
+\label{fig:fpvsoop}
+\end{figure}
+
+What we evince from the gathered data is that, normally, there is no big difference between the LOC for FP and OOP.
+Nevertheless, in certain algorithms the LOC in FP implementations are less than the half of OOP implementation.
 

### 2015-10-10 article.tex
Tables of results and references added

@@
 
-The experimentation consist in static code analysis on a set of eight algorithms, namely:
+The experimentation consist in static code analysis on a set of eight algorithms\cite{algorithms}, namely:
 
@@
 \item n-body
-\item pidigits
+\item pi-digits
 \item regex-dna
@@
 
+\begin{table}[htbp]
+\centering
+\caption{LOC for different implementations.}
+\label{tab:all}
+    \begin{tabular}{rcccccccccc}
+
+ & Haskell & Erlang & Scala & Clojure & F\#  & Python & C++ & C\#  & Go  & Java \\ 
+\hline
+        FASTA                & 111     & 61     & 91    & 129     & 69  & 84     & 132 & 144 & 162 & 292  \\ 
+        k-nucleotide         & 159     & 67     & 194   & 266     & 104 & 156    & 117 & 168 & 175 & 143  \\ 
+        Mandelbrot           & 49      & 42     & 59    & 110     & 54  & 59     & 79  & 63  & 103 & 55   \\ 
+        n-body               & 166     & 114    & 117   & 189     & 98  & 89     & 132 & 100 & 127 & 138  \\ 
+        Pi-digit             & 13      & 33     & 88    & 266     & 89  & 33     & 72  & 113 & 85  & 110  \\ 
+        regex-dna            & 106     & 95     & 148   & 51      & 52  & 38     & 62  & 38  & 77  & 148  \\ 
+        reverse-compl.   & 68      & 136    & 66    & 64      & -   & 61     & 222 & 85  & 174 & 182  \\ 
+        spectral-norm        & 22      & 35     & 50    & 94      & 59  & 39     & 161 & 109 & 72  & 98  \\
+\hline
+    \end{tabular}
+\end{table}
+
+\begin{table}[htbp]
+\centering
+\caption{LOC for the two paradigms.}
+\label{tab:all}
+    \begin{tabular}{rcccccccccc}
+
+ & FP & OOP \\ 
+\hline
+        FASTA  & 91 & 183 \\ 
+        k-nucleotide & 158 & 151\\ 
+        Mandelbrot & 63 & 75 \\ 
+        n-body & 129 & 125 \\ 
+        Pi-digit & 87 & 95 \\
+        regex-dna & 82 & 65 \\ 
+        reverse-compl. & 66 & 166 \\ 
+        spectral-norm & 50 & 110 \\
+\hline
+    \end{tabular}
+\end{table}
+
 An alternative way to visualise these results is to think in terms of language and display which algorithms is the best to implement and be more productive.
@@
 What we evince from the gathered data is that, normally, there is no big difference between the LOC for FP and OOP.
-Nevertheless, in certain algorithms the LOC in FP implementations are less than the half of OOP implementation.
+Nevertheless, in algorithms, like FASTA, revers-complement and spectral-norm, the LOC in FP implementations are less than the half of OOP implementation.
+This find explanation in \cite{fpmatter} where the benefits of FP are discussed.
+Among the most significative there are: the syntax (which is simpler), higher order functions and the standard list-manipulation primitives.
 
 \section{Discussion}
+
+Static code analysis on the lines of code itself is not enough to have a global vision about the two paradigms.
+In fact, there are many additional factors that affect productivity such as the documentation, the comments, the development time, the application domain, the experience of the developer, the time to implement modifications\cite{harrison, Haskell-vs-ada}.
+
+For example, in Fig.
+\ref{fig:hudak} we notice that a very concise implementation may require a bigger documentation, like in the case of (1).
+Furthermore, we observe that in (1), (2), (3), (10) the development time is proportional to the LOC and this fact support our hypothesis.
+
+It has not been possible to succeed in the experimentation concerning the CR as we could not find a suitable case study\cite{play, flink, rx}.
+For instance, many implementations are based on the JVM so that the developers reuse their Java implementation building an API layer for functional languages, like Scala and Clojure.
+
+\begin{figure}[h]
+\includegraphics[width=\textwidth]{4}
+\caption{The result of Hudak and Jones experimentation in\cite{Haskell-vs-ada}.}
+\label{fig:hudak}
+\end{figure}
 

### 2015-10-11 article.tex
Conclusions added and general review.

@@
 Both developers and managers try to maximise their productivity, in order to increase as much as possible the profits or gain market advantage.
-We assume that choosing an appropriate programming paradigm is a critical factor that leads to saving of these resources.
+We assume that choosing an appropriate programming paradigm is a critical factor that leads to a better allocation of human resources.
 
@@
 In section 2, a summary of the studied literature will be presented while in section 3 the methods and the hypothesis will be discussed.
-The results of the investigation and the analysis are presented in section 5, in section 6 a discussion about the initial hypothesis and the results will be discussed.
+The results of the investigation and the analysis are presented in section 5, in section 6 a discussion about the problems and additional work is provided.
 
@@
 
-The experimentation consist in static code analysis on a set of eight algorithms\cite{algorithms}, namely:
+The experimentation consists in static code analysis on a set of eight algorithms\cite{algorithms}, namely:
 
@@
 
-The previous set of algorithms has been implemented in the following functional languages (both pure and not):
+This set of algorithms has been implemented in Haskell, Erlang, Scala, Clojure, F\# and Python for the functional programming paradigm, in C++, C\#, Go and Java for the object-oriented paradigm.
 
-\begin{itemize}
-\item Haskell
-\item Erlang
-\item Scala
-\item Clojure
-\item F\#
-\item Python
-\end{itemize}
-The object-oriented languages these algorithm were implemented with are:
+In the experimentation we analysed the LOC for each language per each algorithm and the results are shown in Table \ref{tab:all}.
+In Fig.
+\ref{fig:alg} is possible to see which language is the most efficient in terms of lines of code per each single algorithm.
 
-\begin{itemize}
-\item C++
-\item C\#
-\item Go
-\item Java
-\end{itemize}
-
-In the experimentation we analysed the LOC for each language per each algorithms and the results are shown in Fig.
-\ref{fig:alg}.
-In this graph it is possible to see which language is the most efficient in terms of lines of code per each single algorithm.
+An alternative way to visualise these results is to think in terms of language and display which algorithms is the best to implement in and to be more productive.
+Fig \ref{fig:lang} shows this alternative representation.
 
@@
 \caption{LOC for the two paradigms.}
-\label{tab:all}
+\label{tab:fpvsoop}
     \begin{tabular}{rcccccccccc}
@@
 
-An alternative way to visualise these results is to think in terms of language and display which algorithms is the best to implement and be more productive.
-Fig \ref{fig:lang} shows this alternative representation.
-
 \begin{figure}[h]
@@
 
-The above graphs have the disadvantages to present too much information so that the readers may have trouble understanding which paradigm is more efficient.
-For this reason we synthesised the gathered data in a graph that shows only the average LOC for each paradigm per each algorithm.
-This representation of the data is shown in Fig.
-\ref{fig:fpvsoop}.
+Table \ref{tab:all} contains many measurements and it is not immediate to understand its meaning.
+Table \ref{tab:fpvsoop} summarise the aforementioned results in terms of FP and OOP, where the average of the LOC are the entries of the table for the two paradigms, and Fig.
+\ref{fig:fpvsoop} is the graphical equivalent representation.
 
@@
 What we evince from the gathered data is that, normally, there is no big difference between the LOC for FP and OOP.
-Nevertheless, in algorithms, like FASTA, revers-complement and spectral-norm, the LOC in FP implementations are less than the half of OOP implementation.
+Nevertheless, in algorithms like FASTA, revers-complement and spectral-norm, the LOC in FP implementations are less than the half of OOP implementation.
 This find explanation in \cite{fpmatter} where the benefits of FP are discussed.
-Among the most significative there are: the syntax (which is simpler), higher order functions and the standard list-manipulation primitives.
+Among the most significative there are: the syntax (which is simpler), higher order functions and the standard list-manipulation primitives (which offer better abstraction and code reuse).
 
 \section{Discussion}
+
+It has not been possible to succeed in the experimentation concerning the CR as we could not find a suitable case study\cite{play, flink, rx}.
+For instance, many implementations are based on the JVM so that the developers reuse their Java implementation building an API layer for functional languages, like Scala and Clojure.
 
@@
 
-For example, in Fig.
-\ref{fig:hudak} we notice that a very concise implementation may require a bigger documentation, like in the case of (1).
-Furthermore, we observe that in (1), (2), (3), (10) the development time is proportional to the LOC and this fact support our hypothesis.
-
-It has not been possible to succeed in the experimentation concerning the CR as we could not find a suitable case study\cite{play, flink, rx}.
-For instance, many implementations are based on the JVM so that the developers reuse their Java implementation building an API layer for functional languages, like Scala and Clojure.
+Nevertheless, we strongly believe that the LOC are directly involved in the productivity and our hypothesis finds confirmation in \cite{Haskell-vs-ada}.
+In fact, in Fig.
+\ref{fig:hudak} we notice that in (1), (2), (3), (10) the development time is proportional to the LOC and this fact support our hypothesis.
 
@@
 \section{Conclusion}
+
+In this investigation we wanted to see whether the functional programming paradigm would affect positively the productivity and the code reuse during the software development process.
+
+Unfortunately, we were not able to check whether it is easier for functional programmers to upgrade software in new releases due to difficulties in finding an appropriate case study.
+In fact, many framework has been taken as case study but in none of them we could find a parallel branch.
+Usually, these framework has been early developed in Java and thanks to the JVM they could reuse the already existing Java code for feeding an API layer in other JVM-based functional languages.
+
+We were able to correlate the productivity through the analysis of the LOC and the development time.
+After the evaluation we can state that FP has real advantages over OOP as it is usually less verbose and this implies time saving.
+
+Even though we could not demonstrate benefits in CR, our findings suggest that the functional programming paradigm might really help developers in terms of productivity in order to improve the allocation of human resources.
 

### 2015-10-23 article.tex
changes

@@
 
-\title{Functional vs. Object Oriented Programming}
+\title{Functional vs. Object Oriented Programming: Productivity comparison}
 \author{Lorenzo Corneo and Nikolaos-Ektoras Anestos \\
@@
 From the evaluation of the gathered data it is possible to see that, averagely, the number of Lines Of Code in a functional programming language is smaller than in an Object Oriented Programming language.
-As a consequence, Functional Programming allows an increased productivity (fewer lines of code correspond to a time saving).
+As a consequence, Functional Programming allows an increased productivity (fewer lines of code correspond to a time saving during development).
 \end{abstract}
@@
 \begin{itemize}
-\item FASTA
-\item k-nucleotide
-\item Mandelbrot
-\item n-body
-\item pi-digits
-\item regex-dna
-\item reverse-complement
-\item spectral-norm
+\item FASTA, generate and write random DNA sequences
+\item k-nucleotide, repeatedly update hashtables and k-nucleotide strings
+\item Mandelbrot, generate a Mandelbrot set and write a portable bitmap
+\item n-body, perform an N-body simulation of the Jovian planets
+\item pi-digits, calculate the digits of Pi with arbitrary-precision arithmetic
+\item regex-dna, match DNA 8-mers and substitute nucleotides for IUB code
+\item reverse-complement, read DNA sequences and write their reverse-complement
+\item spectral-norm, calculate an eigenvalue using the power method
 \end{itemize}

### 2015-10-23 article.tex
Newer version

@@
 
-A quantitative method is used for this project because the authors will gather data from many sources and they will report in the paper.
-Afterward, the analysis of the gathered data will be performed so that the initial hypothesis will be either accepted or rejected.
+A quantitative method is used for this project because the authors have gathered data from many sources and they report in this paper.
+Afterward, the analysis of the gathered data is performed so that the initial hypothesis will be either accepted or rejected.
 
-The approach used will be deductive because we started this research with our own hypothesis on the basis of existing theories, for instance, regarding the benefits of FP and the advantages that developers can gain in software production.
+The approach used is deductive because we started this research with our own hypothesis on the basis of existing theories, for instance, regarding the benefits of FP and the advantages that developers can gain in software production.
 
-Afterward, when the data will be acquired, the authors will perform observation and analysis of the gotten results. From that, it will be possible to accept or reject the initial hypothesis.
+Afterward, when the data has been acquired, the authors perform observation and analysis of the gotten results. From that, it will be possible to accept or reject the initial hypothesis.
 
@@
 This set of algorithms has been implemented in Haskell, Erlang, Scala, Clojure, F\# and Python for the functional programming paradigm, in C++, C\#, Go and Java for the object-oriented paradigm.
+
+\subsection{Lines of Code}
 
@@
 
-\section{Discussion}
+\subsection{Code Reuse}
 
@@
 For instance, many implementations are based on the JVM so that the developers reuse their Java implementation building an API layer for functional languages, like Scala and Clojure.
+
+\section{Discussion}
 

### 2015-10-23 article.tex
method changed

@@
 Afterward, when the data has been acquired, the authors perform observation and analysis of the gotten results. From that, it will be possible to accept or reject the initial hypothesis.
+
+Static code analysis on the set of algorithms described below has been performed using the command line tool called Count Lines of Code (cloc) \cite{cloc}, which counts lines of code, blank lines and comments in multiple languages.
+There are several options on how to present the data using this tool, i.e.
+one can obtain an analysis of a language used inside a project (several files), or extract results for each file individuals etc.
 

### 2015-10-25 article.tex
some fixing

--- 2015-10-23 article.tex
@@
 \begin{itemize}
-\item FASTA, generate and write random DNA sequences
-\item k-nucleotide, repeatedly update hashtables and k-nucleotide strings
-\item Mandelbrot, generate a Mandelbrot set and write a portable bitmap
-\item n-body, perform an N-body simulation of the Jovian planets
-\item pi-digits, calculate the digits of Pi with arbitrary-precision arithmetic
-\item regex-dna, match DNA 8-mers and substitute nucleotides for IUB code
-\item reverse-complement, read DNA sequences and write their reverse-complement
-\item spectral-norm, calculate an eigenvalue using the power method
+\item FASTA, generate and write random DNA sequences.
+\item k-nucleotide, repeatedly update hashtables and k-nucleotide strings.
+\item Mandelbrot, generate a Mandelbrot set and write a portable bitmap.
+\item n-body, perform an N-body simulation of the Jovian planets.
+\item pi-digits, calculate the digits of Pi with arbitrary-precision arithmetic.
+\item regex-dna, match DNA 8-mers and substitute nucleotides for IUB code.
+\item reverse-complement, read DNA sequences and write their reverse-complement.
+\item spectral-norm, calculate an eigenvalue using the power method.
 \end{itemize}

### 2015-10-31 article.tex
Future work added

@@
 
+\section{Future work}
+
+We have just demonstrated with a quantitative approach that FP paradigm has real benefits that help the developers to increase their productivity when writing code.
+Then, a question raise spontaneously: why FP languages are so unpopular, accordingly to TIOBE Index (October 2015)\cite{tiobe}, if they have all of these benefits?
+
+We strongly believe that the most important factor that cuts off from the market these languages is the lack of a consistent and agile Graphic User Interface library support.
+Nowadays, the user interface and the user experience have become an essential requirement for every kind software.
+Furthermore, many object oriented implementation, in complex software systems, are more verbose than functional programming because of the code concerning the GUI building.
+It is not a coincidence that Erlang, for example, runs in the backend of Facebook, WhatsApp and many other software systems.
+
+For future work, when also FP languages will have a valuable GUI support, we want to consider complex systems (not only algorithms) and investigate how the new results will differ from the ones presented here.
+To conclude, as already mentioned in the previous section, in order to have a global view on the topic we want to analyse other meaningful parameters (i.e.
+documentation, the comments, the development time, the application domain, etc..) and see if FP paradigm keeps on being better than OOP paradigm.
+
 \section{Conclusion}

### 2015-10-31 article.tex
final?

--- 2015-10-25 article.tex
@@
 Human resources and time consumption are very important factors in software development.
-Both developers and managers try to maximise their productivity, in order to increase as much as possible the profits or gain market advantage.
-We assume that choosing an appropriate programming paradigm is a critical factor that leads to a better allocation of human resources.
+Both developers and managers try to maximize their productivity, in order to increase as much as possible the profits, or gain market advantage.
+We assume that choosing an appropriate programming paradigm is a critical factor that leads to a better allocation of human resources during software development.
 
@@
 
-Functional programming has some fundamental benefits that increase programmer productivity over that when OOP languages.
-The purpose of this project is to demonstrate some of the benefits of functional programming over OOP, in order to persuade developers to switch to functional programming.
+Functional Programming languages, have some fundamental benefits that increase programmer productivity over Object Oriented Programming languages.
+The purpose of this project is to demonstrate some of the benefits of functional programming over OOP, in order to persuade developers to adopt functional programming and use it in more and more cases.
 
@@
 The paper presents the results of a quantitative evaluation of code implemented in both Functional Programming and Object Oriented Programming.
-The main parameters that will be analysed are the Line Of Code and the Code Reuse.
+The main parameters that will be analyzed are the Line Of Code and the Code Reuse.
 
@@
 The CR is evaluated counting the difference in terms of LOC between adjacent releases of a program.
-Object of the investigation, as a case study, is the asynchronous web server Play framework\cite{play}, that is implemented both in Java and Scala.
+Object of the investigation, as a case study, are the asynchronous web server Play framework\cite{play}, apache Flink\cite{flink} and Reactive X library\cite{rx} that are implemented both in Java and Scala.
 
 In section 2, a summary of the studied literature will be presented while in section 3 the methods and the hypothesis will be discussed.
-The results of the investigation and the analysis are presented in section 5, in section 6 a discussion about the problems and additional work is provided.
+The results of the investigation and the analysis are presented in section 5, in section 6 a discussion about the problems and finally we recognize the future work that needs to be done in section 7.
 
@@
 OOP is the de facto standard programming paradigm for companies and widely adopted all around the world.
-This find a confirmation in the TIOBE Index\cite{tiobe}.
+This finds confirmation in the TIOBE Index\cite{tiobe}.
 In fact, in the first positions only OOP and Web programming languages are found, while the first FP language is Scala at the 24th position with a very low rating of 0.781\%.
@@
 
-Human resources and time consumption are very important factors in software development, especially for companies that want to optimise profits and gain advantages from the market.
+Human resources and time consumption are very important factors in software development, especially for companies that want to optimize profits and gain advantages over the market.
 
-We strongly believe that, using functional programming languages, developers will be more productive and they will also produce higher quality code that will allow them to reuse much of it when releasing newer versions of their programs.
+We strongly believe that, using functional programming languages, developers will be more productive and they will also produce higher quality code that will allow them to reuse much of it when updating their programs.
 
@@
 
-The approach used is deductive because we started this research with our own hypothesis on the basis of existing theories, for instance, regarding the benefits of FP and the advantages that developers can gain in software production.
+The approach used is deductive.
+The authors started this research stating their hypothesis on the basis of existing theories, for instance, regarding the benefits of FP and the advantages that developers can gain in software production.
 
-Afterward, when the data has been acquired, the authors perform observation and analysis of the gotten results. From that, it will be possible to accept or reject the initial hypothesis.
+Afterwards, when the data has been acquired, the authors perform observation and analysis of the gathered results. From that, it will be possible to accept or reject the initial hypothesis.
 
@@
 \item FASTA, generate and write random DNA sequences.
-\item k-nucleotide, repeatedly update hashtables and k-nucleotide strings.
+\item k-nucleotide, repeatedly update hash tables and k-nucleotide strings.
 \item Mandelbrot, generate a Mandelbrot set and write a portable bitmap.
@@
 
-In the experimentation we analysed the LOC for each language per each algorithm and the results are shown in Table \ref{tab:all}.
+In the experimentation we analyzed the LOC for each language per each algorithm and the results are shown in Table \ref{tab:all}.
 In Fig.
@@
 
-An alternative way to visualise these results is to think in terms of language and display which algorithms is the best to implement in and to be more productive.
+An alternative way to visualize these results is to think in terms of language and display which algorithms is the best to implement in and to be more productive.
 Fig \ref{fig:lang} shows this alternative representation.
@@
 Table \ref{tab:all} contains many measurements and it is not immediate to understand its meaning.
-Table \ref{tab:fpvsoop} summarise the aforementioned results in terms of FP and OOP, where the average of the LOC are the entries of the table for the two paradigms, and Fig.
+Table \ref{tab:fpvsoop} summarizes the aforementioned results in terms of FP and OOP, where the average of the LOC are the entries of the table for the two paradigms, and Fig.
 \ref{fig:fpvsoop} is the graphical equivalent representation.
@@
 This find explanation in \cite{fpmatter} where the benefits of FP are discussed.
-Among the most significative there are: the syntax (which is simpler), higher order functions and the standard list-manipulation primitives (which offer better abstraction and code reuse).
+Some of the most significant are: the syntax (which is simpler), higher order functions and the standard list-manipulation primitives (which offer better abstraction and code reuse).
 
@@
 For instance, many implementations are based on the JVM so that the developers reuse their Java implementation building an API layer for functional languages, like Scala and Clojure.
+
+Also, in most cases, the features that the developers provide in each version are not exactly the same for the two paradigms.
+For example version 2.0 has features X, Y, Z in OOP but only feature X is implemented for the functional API.
+This makes it impossible to compare the two paradigms over this parameters, but we found that in the case where a feature is present for both paradigms, the FP API needs always less lines of code to call it, over the OOP counterpart.
 
@@
 
+\section{Future work}
+
+We have just demonstrated with a quantitative approach that FP paradigm has real benefits that help the developers increase their productivity when writing code.
+Then, a question raises spontaneously: why FP languages are so unpopular, accordingly to TIOBE Index (October 2015)\cite{tiobe}, if they have all of these benefits?
+
+We strongly believe that the most important factor that cuts off from the market these languages is the lack of a consistent and agile Graphic User Interface library support.
+Nowadays, the user interface and the user experience have become an essential requirement for every kind software.
+Furthermore, many object oriented implementations, in complex software systems, are more verbose than functional programming because of the code concerning the GUI building.
+It is not a coincidence that Erlang, for example, runs in the back-end of Facebook, WhatsApp and many other software systems.
+
+For future work, when also FP languages will have a valuable GUI support, we want to consider complex systems (not only algorithms) and investigate how the new results will differ from the ones presented here.
+As already mentioned in the previous section, in order to have a global view on the topic we want to analyze other meaningful parameters (i.e.
+documentation, the comments, the development time, the application domain, etc..) and see if FP paradigm keeps on being better than OOP paradigm.
+
+Finally, one other aspect that managers and companies need to take into consideration when they want to hire programmers, is their cost.
+At the moment, the wide adoption of OOP languages over FP makes functional programmers less common which means that they are also more expensive.
+Our study so far, tries to show the advantages of FP in order to make new developers adopt it, reducing the aforementioned gap.
+
 \section{Conclusion}
@@
 Unfortunately, we were not able to check whether it is easier for functional programmers to upgrade software in new releases due to difficulties in finding an appropriate case study.
-In fact, many framework has been taken as case study but in none of them we could find a parallel branch.
-Usually, these framework has been early developed in Java and thanks to the JVM they could reuse the already existing Java code for feeding an API layer in other JVM-based functional languages.
+In fact, many frameworks has been taken as case studies but in none of them we could find a parallel branch.
+Usually, these framework has been early developed in Java and thanks to the JVM they could reuse the already existing Java code to feed an API layer in other JVM-based functional languages.
 
 We were able to correlate the productivity through the analysis of the LOC and the development time.
-After the evaluation we can state that FP has real advantages over OOP as it is usually less verbose and this implies time saving.
+After the evaluation we can state that FP has real advantages over OOP as it is usually less verbose and this implies time saving during development.
 
